Modelling last glacial cycle ice dynamics in the Alps

### 2017-07-06 articles/cycle/alps-cycle.tex
Update parameter table to alpcyc4 config.
Add subsection on basal topography.

\section{Model setup}
\label{sec:model}

@@

      The simulations presented here were run using the Parallel Ice Sheet
      Model (PISM, version~0.7.2), an open source, finite
      difference, shallow ice sheet model \citep{PISM-authors.2015}. The
      model requires input on basal topography, sea level, geothermal heat
      flux and climate forcing. It computes the evolution of ice extent and
      thickness over time, the thermal and dynamic states of the ice sheet,
      and the associated lithospheric response.

      Basal topography is bilinearly interpolated from the ETOPO1 combined
      topography and bathymetry dataset with a~resolution of 1\,arc-min
      \citep{Amante.Eakins.2009} to the model grids. It responds to ice load
      following a~bedrock deformation model that includes local isostasy,
      elastic lithosphere flexure and viscous astenosphere deformation in an
      infinite half-space \citep{Lingle.Clark.1985,Bueler.etal.2007}.
      A~relatively low viscosity value of $\nu_{\mathrm{m}} =
      1\times10^{19}$\,\unit{Pa\,s} is used for the astenosphere
      (Table~\ref{tab:params}) in accordance with the results from regional
      glacial isostatic adjustment modelling at the northern Cascadia
      subduction zone \citep{James.etal.2009}. Sea level is lowered as
      a~function of time based on the Spectral Mapping Project
      \citep[SPECMAP,][]{Imbrie.etal.1989} time scale.

@@

@@

        $E_{\text{SIA}}$   & SIA enhancement factor
                & 5
                & --
                & \citet{Cuffey.Paterson.2010} \\

@@

        $\nu_{\mathrm{m}}$ & Astenosphere viscosity
                & $1\e{19}$
                & \unit{Pa\,s}
                & \citet{James.etal.2009} \\

        $\rho_{\mathrm{l}}$& Lithosphere density
                & 3300
                & \unit{kg\,m^{-3}}
                & \citet{Lingle.Clark.1985} \\

        $D$     & Lithosphere flexural rigidity
                & $5\e{24}$
                & \unit{N\,m}
                & \citet{Lingle.Clark.1985} \\

@@

>>>>

@@

    We use the Parallel Ice Sheet
    Model (PISM, development version~e9d2d1f), an open source, finite
    difference, shallow ice sheet model \citep{PISM-authors.2017}. The
    model requires input on basal topography, geothermal heat
    flux and climate forcing. It computes the evolution of ice extent and
    thickness over time, the thermal and dynamic states of the ice sheet,
    and the associated lithospheric response.

@@

\subsection{Basal topography}
\label{sec:bedrock}

    The initial basal topography is bilinearly interpolated from the
    hole-filled, Shuttle Radar Topography Mission (SRTM) data with a~resolution
    of 30\,arc-sec \citep{Jarvis.etal.2008}. These data include post-glacial
    sediment fills and lakes surface topography. However, they were corrected
    for an estimation of present-day ice thickness according from modern
    glacier outlines, surface topography and simplified ice physics
    \citep{Huss.Farinotti.2012}.

    Basal topography responds to ice load following a~bedrock deformation model
    that includes local isostasy, elastic lithosphere flexure and viscous
    astenosphere deformation in an infinite half-space
    \citep{Lingle.Clark.1985,Bueler.etal.2007}. The astenosphere viscosity,
    $\nu_{\mathrm{m}}=2.2\times10^{20}$\,\unit{Pa\,s}, the astenosphere
    density, $\rho_{\mathrm{m}}=3300$\,\unit{kg\,m^{-3}}, and the lithosphere
    elastic rigidity, $D=1.389\e{24}$\,\unit{N\,m}, were set according to
    results from glacial isostatic adjustment modelling of deglacial rebound in
    the Alps most closely reproducing observed modern uplift rates
    \citep[Table~\ref{tab:params};][Supplementary Fig.~7]{Mey.etal.2016}. The
    latter was computed as $D=YE^3/12(1-\nu)$, where $Y=100$\,GPa is Young's
    modulus, $\nu=0.25$ is the Poisson ratio, and $E=50$\,km is the average
    effective elastic thickness of the lithosphere
    \citep[Table~\ref{tab:params};][]{Mey.etal.2016}.

@@

        $E_{\text{SIA}}$   & SIA enhancement factor
                & 2
                & --
                & \citet{Cuffey.Paterson.2010} \\

@@

        $\nu_{\mathrm{m}}$ & Astenosphere viscosity
                & $2.2\e{20}$
                & \unit{Pa\,s}
                & \citet{Mey.etal.2016} \\

        $\rho_{\mathrm{m}}$& Astenosphere density
                & 3300
                & \unit{kg\,m^{-3}}
                & \citet{Mey.etal.2016} \\

        $D$     & Lithosphere flexural rigidity
                & $1.389\e{24}$
                & \unit{N\,m}
                & \citet{Mey.etal.2016} \\

@@

        $\omega$& Precipitation factor
                & 0.0704
                & --
                & \citet{Huybrechts.2002} \\
